---
title: "Nodiv results"
author: "Marisa Lim"
date: "December 9, 2015"
output: 
  html_document:
    theme: flatly
---

Run nodiv on hummingbird species range map data (site x sp matrix) and climate data

```{r, eval = TRUE, echo = FALSE, include = FALSE}
# load libraries
mypackages <- c('devtools', 'ape', 'raster', 'fields', 'dplyr', 'picante')
lapply(mypackages, require, character.only=T)
install_github("mkborregaard/nodiv") #Gives the newest version - is probably LESS buggy than cran version

# load this for the gridData function
load('C:/Users/mcwlim/Desktop/Github/Hbird-diversity/P3_Diversi/nodiv-trait-analysis.RDATA')
```

#### 1. Load input data: sitexsp, coords, shape, tree 
```{r, eval = TRUE, echo = FALSE, include = FALSE}
# Read in site by species matrix
#hb_commat <- read.csv("for sitexsp/hbird_sitexspmat_sub.csv") #a 0/1 site-by-species matrix
hb_commat <- read.csv("C:/Users/mcwlim/Dropbox/resultsfromSarahcomputer/nodesigfiles/hbird_sitexspmat_sub.csv")
dim(hb_commat)
hb_commat[1:5,1:4]
rownames(hb_commat) <- hb_commat$X
hb_commat2 <- hb_commat[,-1]

# Read in cell coordinate information
#hb_coords <- read.csv("for sitexsp/cell_coords_rounded.csv")[,-1] # a data.frame with one row for each site, and lat/long coordinates in 2 of the columns
hb_coords <- read.csv("C:/Users/mcwlim/Dropbox/resultsfromSarahcomputer/nodesigfiles/cell_coords_rounded.csv")[,-1]
head(hb_coords)
colnames(hb_coords) <- c("cellIDs", "Long", "Lat")

# Read in geographic extent shape file
#myshape <- raster("for sitexsp/myvarscrop.grd")
myshape<- raster("C:/Users/mcwlim/Dropbox/resultsfromSarahcomputer/nodesigfiles/myvarscrop.grd")

# Read in tree
#tree <- read.tree("humtree280.tre") 
tree <- read.tree("C:/Users/mcwlim/Desktop/StonyBrook/GrahamLab/Dissertation idea materials/THESIS PROJECTS/Hummingbird_Genomics/humtree280.tre")
```

#### 2. Load inputs into nodiv object 
```{r, eval = TRUE, echo = FALSE, include = FALSE}
# Get environmental variables
# NOTE: r script version has altitude, i removed it for this doc because it doesn't add too much
# # Get rasters res=10
tmin <- raster::getData("worldclim", var="tmin", res = 10) %>% mean()
prec <- raster::getData("worldclim", var="prec", res = 10) %>% sum()
# Extract values (res=10)
hb_coords$tmin <- raster::extract(x=tmin, y=hb_coords[c("Long", "Lat")])/10
hb_coords$tmin[is.na(hb_coords$tmin)] <- na.omit(hb_coords$tmin)/10
hb_coords$prec <- raster::extract(x=prec, y=hb_coords[c("Long", "Lat")])
hb_coords$prec[is.na(hb_coords$prec)] <- na.omit(hb_coords$prec)
hb_coords2 <- arrange(hb_coords, cellIDs)
head(hb_coords2)

# Generate distrib_data object
hummers <- distrib_data(hb_commat2, hb_coords) %>% 
  add_shape(myshape)
# res=10
hummers <- add_sitestat(distrib_data=hummers, 
                        sitestat=dplyr::select(hb_coords2, tmin, prec), 
                        site=rownames(hb_commat2))
# Add phylogeny
hummers <- nodiv_data(tree, hummers) 

# Grid Data...
#...in geographic space
hummers_geo <- gridData(hummers, type = "geo") 
#...in environmental space 
hummers_env <- gridData(hummers, type = "env", env_binsizes = c(2, 400))
  ## Warning message: In points2grid(points, tolerance, round) :   grid has empty column/rows in dimension 2
```

#### The most basic summary metrics are then available:
Here, showing species richness at several different nodes in the phylogeny:
```{r, echo=FALSE, fig.align='center'}
plot_richness(hummers, main="Species richness: full tree")

par(mfrow=c(1,2))
plot_node(hummers, node=329, main="Species richness: \nAndean clade vs. rest")
plot_node(hummers, node=272, main="Species richness \nat node: Topazes,Hermits vs. rest")


plot_node(hummers,node=416, main="Species richness:\n out of SA")
plot_node(hummers,node=330, main="Species richness:\n Coquettes vs. Brilliants")
```

#### 3. Run nodiv: geographic space 
```{r, eval = TRUE, echo = FALSE, include = FALSE}
humm_nodiv <- Node_analysis(hummers, repeats = 100, method = "rdtable") 
```
```{r, eval = TRUE}
summary(humm_nodiv)
```

Results for geographic space analysis:
GND values plotted at nodes
```{r, echo=FALSE, fig.align='center'}
plot(humm_nodiv, label=humm_nodiv$GND, col=heat.colors(10), direction="upwards")
```

SOS values plotted for several nodes:
```{r, echo=FALSE, fig.align='center'}
#plot SOS values for a given node
par(mfrow=c(1,3))
plotSOS(humm_nodiv, 329, main="Andean clade vs. rest", col="HMrainbow", xlab="Longitude", ylab="Latitude") 

plotSOS(humm_nodiv, 272, main="HermitsTopazes vs. rest", col="HMrainbow", xlab="Longitude", ylab="Latitude")

plotSOS(humm_nodiv, 416, main="BeesGems vs. Emeralds", col="HMrainbow", xlab="Longitude", ylab="Latitude")
```

#### 4. Run nodiv: environmental space 
```{r, eval = TRUE, echo = FALSE, include = FALSE}
humm_nodiv_env <- Node_analysis(hummers_env, repeats = 100, method = "rdtable")
```

Results for geographic space analysis:
GND values plotted at nodes
```{r, echo=FALSE, fig.align='center'}
plot(humm_nodiv_env, label=humm_nodiv_env$GND, col=heat.colors(10), direction="upwards")
```

SOS values plotted for several nodes:

* x-axis = temperature (res=10 arcminutes, 0.167 degrees)

* y-axis = precipitation (res=10 arcminutes, 0.167 degrees)

1. Base node of phylogeny vs. rest
```{r, echo=FALSE, fig.align='center'}
plotSOS(humm_nodiv_env, 272, col=HMrainbow()) 
```

2. Andean clade vs. rest
```{r, echo=FALSE, fig.align='center'}
plotSOS(humm_nodiv_env, 329, col=HMrainbow())
```

3. Base of Coquette split
```{r, echo=FALSE, fig.align='center'}
plotSOS(humm_nodiv_env, 372, col=HMrainbow())
```

4. Bees and Mt Gems vs. Emeralds
```{r, echo=FALSE, fig.align='center'}
plotSOS(humm_nodiv_env, 416, col=HMrainbow())
```



